Ethereum: A Secure Decentralised Generalised Transaction Ledger

### 2014-04-04 Paper.tex
Most of instruction set documented.

\documentclass[9pt,oneside]{amsart}
%\usepackage{tweaklist}
\usepackage{url}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[a4paper,width=170mm,top=18mm,bottom=22mm,includeheadfoot]{geometry}
\usepackage{booktabs}
\usepackage{array}
\usepackage{verbatim}
\usepackage{caption}
\usepackage{natbib}
\usepackage{float}
\usepackage{pdflscape}
\usepackage{mathtools}

@@

\title{Ethereum: A Secure Decentralised Transaction Progression Platform}
\author{
    Gavin Wood\\
    CTO, Ethereum Project\\
    gavin@ethereum.org
}
\begin{document}

@@

\begin{equation}
H_d :=  \begin{cases}
2^{22} & \quad \text{if genesis block}\\
H'_d + \frac{H'_d}{1024} & \quad \text{if $H_t < H'_t + 42$}\\
H'_d - \frac{H'_d}{1024} & \quad \text{otherwise}\\
\end{cases}
\end{equation}

@@

\begin{eqnarray}
a' & \equiv & \mathcal{L}_{160}(\mathcal{S}(\mathcal{R}([s, o])))\\
a & \equiv & \arg \min_x : x \geq a' \wedge \mathbb{S}[x] = \emptyset
\end{eqnarray}

where $\mathcal{S}$ is the SHA3 256-bit hash function, $\mathcal{R}$ is the RLP encoding function, $\mathcal{L}_y(X)$ evaluates to the leftmost $y$ bits of the binary data $X$ and $\mathbb{S}[x]$ is the address state of $x$ or $\emptyset$ if none exists.

@@

\begin{equation}
\mathbb{S}'[r] := ( 0, v, \empty, \mathcal{S}(\mathbf{b}) )
\end{equation}

@@

\begin{equation}
\mathbb{S}'[r]_{balance} := \mathbb{S}[r]_{balance} + v
\end{equation}

@@

\begin{itemize}
\item $I_a$, the address of the account which owns the code that is executing.
\item $I_o$, the sender address of the transaction that originated this execution.
\item $I_p$, the price of gas in the transaction that originated this execution.
\item $I_D$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
\item $I_v$, the value passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
\item $I_H$, the block header of the previous block.
\item $I_{H'}$, the block header of the current block.
\item $I_n$, the block number of the current block.
\item $I_C$, the byte array that is the machine code to be executed.
\end{itemize}

@@

Certain instructions do; alter them---such changes are noted in Appendix \ref{app:vm}.

\subsubsection{Terms}

Summary line:
<Op-code>: <Mnemonic> -<R> +<A>

If PC is not defined explicitly, then it must be assumed PC := PC' + 1. Exceptions are .

The cardinality of T (i.e. size of the stack) is altered by A - R between T \& T', by adding or removing items as necessary from the front.

Where:
R: The minimal cardinality of the stack for this instruction to proceed. If this is not achieved then the machine halts with a stack underflow exception. (Note: In some cases of some implementations, this is also the number of values "popped" from the implementation's stack during the course of instruction execution.)
(A - R): The net change in cardinality of the stack over the course of instruction execution.

$\forall$ i: if T[i] is not defined explicitly, then it must be assumed T[i] := T'[i + R - A] where i + R >= A.
$\forall$ i: if M[i] is not defined explicitly, then it must be assumed M[i] := M'[i].
$\forall$ i: if S[i] is not defined explicitly, then it must be assumed S[i] := S'[i].

The expression (COND ? ONE : ZERO), where COND is an expression and ONE and ZERO are both value placeholders, evaluates to ONE if COND is true, and ZERO if COND is false. This is similar to the C-style ternary operator.

When a 32-byte machine datum is interpreted as a 160-bit address or hash, the rightwards 20 bytes are taken (i.e. the low-order bytes when interpreting the data as Big-Endian).

++ is the concatenation operator; all operands are byte arrays (mostly 32-byte arrays here, since that's the size of the VM's data \& address types).

LEFT\_BYTES(A, n) returns the array bytes comprising the first (leftmost) n bytes of the 32 byte array A, which can be considered equivalent to a single value in the VM.

@@

Two directions exist for ASIC resistance; firstly make it sequential memory-hard, i.e. engineer the function such that the determination of the nonce requires a lot of memory and that the memory cannot be used in parallel to discover multiple nonces simultaneously. The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.

@@

\begin{equation}
\mathcal{P}(H, n) \equiv \mathcal{B}(\mathcal{S}(\mathcal{S}(\mathcal{R}(H)) \bigoplus n))
\end{equation}

where:
$\mathcal{R}(H)$ is the RLP encoding of the $H$;
$\mathcal{S}$ is the SHA3 hash function accepting an arbitrary length series of bytes and evaluating to a series of 32 bytes (i.e. 256-bit);
$n$ is the nonce, a series of 32 bytes;
$\bigoplus$ is the series concatenation operator;
$\mathcal{B}(X)$ evaluates to the value equal to X when interpreted as a big-endian-encoded integer.

@@

For the release series, we use a more complex proof-of-work. This has yet to be formally defined, but involves two components; firstly,

@@

\section{Discussion} \label{ch:discussion}

\subsection{Examples} \label{ch:examples}

@@

\subsection{Further Work} \label{ch:further}

@@

@@

0s: arithmetic operations

0x00: STOP -0 +0
Halts execution.
Any gas left over gets returned to caller (or in the case of the top-level call, the sender converted back to ETH).
0x01: ADD -2 +1
S[0] := S'[0] + S'[1]
0x02: MUL -2 +1
S[0] := S'[0] * S'[1]
0x03: SUB -2 +1
S[0] := S'[0] - S'[1]
0x04: DIV -2 +1
S[0] := S'[0] / S'[1]
0x05: SDIV -2 +1
S[0] := S'[0] / S'[1]
S'[0] \& S'[1] are interpreted as signed 256-bit values for the purposes of this operation.
0x06: MOD -2 +1
S[0] := S'[0] % S'[1]
0x07: SMOD -2 +1
S[0] := S'[0] % S'[1]
S'[0] \& S'[1] are interpreted as signed 256-bit values for the purposes of this operation.
0x08: EXP -2 +1
S[0] := S'[0] + S'[1]
0x09: NEG -1 +1
S[0] := -S'[0]
0x0a: LT -2 +1
S[0] := S'[0] < S'[1] ? 1 : 0
0x0b: GT -2 +1
S[0] := S'[0] > S'[1] ? 1 : 0
0x0c: EQ -2 +1
S[0] := S'[0] == S'[1] ? 1 : 0
0x0d: NOT -1 +1
S[0] := S'[0] == 0 ? 1 : 0

10s: bit operations

0x10: AND -2 +1
S[0] := S'[0] AND S'[1]
0x11: OR -2 +1
S[0] := S'[0] OR S'[1]
0x12: XOR -2 +1
S[0] := S'[0] XOR S'[1]
0x13: BYTE -2 +1
S[0] := S'[0]th byte of S'[1]
if S'[0] < 32
S[0] := 0
otherwise
for Xth byte, we count from left - 0th is therefore the leftmost (most significant in BE) byte.

20s: crypto opcodes

0x20: SHA3 -2 +1
S[0] := SHA3( T'[ S'[0] ++ ... ++ (S'[0] + S'[1]) ])

30s: closure state opcodes

0x30: ADDRESS -0 +1
S[0] := ADDRESS
i.e. the address of this closure.
0x31: BALANCE -1 +1
S[0] := B[ S'[0] ]
i.e. the balance of this closure.
0x32: ORIGIN -0 +1
S[0] := A
Where A is the address of the account that made the original transaction leading to the current closure and is paying the fees
0x33: CALLER -0 +1
S[0] := A
Where A is the address of the object that made this call.
0x34: CALLVALUE -0 +1
S[0] := V
Where V is the value attached to this call.
0x35: CALLDATALOAD -1 +0
S[0] := D[ S'[0] ... (S'[0] + 31) ]
Where D is the data attached to this call (as a byte array).
Any bytes that are out of bounds of the data are defined as zero.
0x36: CALLDATASIZE -0 +1
S[0] := DS
Where DS is the number of bytes of data attached to this call.
0x37: GASPRICE -0 +1
S[0] := V
Where V is the current gas price (né fee multiplier).

40s: block operations

0x40: PREVHASH -0 +1
S[0] := H
Where H is the SHA3 hash of the previous block.
0x41: COINBASE -0 +1
S[0] := A
Where A is the coinbase address of the current block.
0x42: TIMESTAMP -0 +1
S[0] := T
Where T is the timestamp of the current block (given as the Unix time\_t when this block began its existence).
0x43: NUMBER -0 +1
S[0] := N
Where N is the block number of the current block (counting upwards from genesis block which has N == 0).
0x44: DIFFICULTY -0 +1
S[0] := D
Where D is the difficulty of the current block.
0x45: GASLIMIT -0 +1
S[0] := L
Where L is the total gas limit of the current block. Always $10^6$.

50s: stack, memory, storage and execution path operations

0x51: POP -1 +0
0x52: DUP -1 +2
S[0] := S'[0]
0x53: SWAP -2 +2
S[0] := S'[1]
S[1] := S'[0]
0x54: MLOAD -1 +1
S[0] := T'[ S'[0] ... S'[0] + 31 ]
0x55: MSTORE -2 +0
T[ S'[0] ... S'[0] + 31 ] := S'[1]
0x56: MSTORE8 -2 +0
T[ S'[0] ... S'[0] + 31 ] := S'[1] \& 0xff
0x57: SLOAD -1 +1
S[0] := P'[ S'[0] ]
0x58: SSTORE -2 +0
P[ S'[0] ] := S'[1]
0x59: JUMP -1 +0
PC := S'[0]
0x5a: JUMPI -2 +0
PC := S'[1] == 0 ? PC' : S'[0]
0x:5b PC -0 +1
S[0] := PC
0x5c: MSIZE -0 +1
S[0] = sizeof(T)
0x5d: GAS
S[0] := G
Where G is the amount of gas remaining after executing the opcode.

60s \& 70s: push

0x60: PUSH1 0 +1
S[0] := C[ PC' + 1 ]
PC := PC' + 2
0x61: PUSH2 0 +1
S[0] := C[ PC' + 1 ] ++ C[ PC' + 2 ]
PC := PC' + 3
...
0x7f: PUSH32 0 +1
S[0] := C[ PC' + 1 ] ++ C[ PC' + 2 ] ++ ... ++ C[ PC' + 32 ]
PC := PC' + 33

f0s: closure-level operations

0xf0: CREATE -5 +1
Immediately creates a contract where:
The endowment is given by S'[0]
The body code of the eventual closure is given by T'[ S'[1] ... ( S'[1] + S'[2] - 1 ) ]
The initialisation code of the eventual closure is given by T'[ S'[3] ... ( S'[3] + S'[4] - 1 ) ]
(Thus the total number of bytes of the transaction data is given by S'[2] + S'[4].)
S[0] = A
where A is the address of the created contract or 0 if the creation failed.
Fees are deducted from the sender balance to pay for enough gas to complete the operation (i.e. contract creation fee + initial storage fee). If there was not enough gas to complete the operation, then all gas will be deducted and the operation fails.
0xf1: CALL -7 +1
Immediately executes a call where:
The recipient is given by S'[0], when interpreted as an address.
The value is given by S'[1]
The gas is given by S'[2]
The input data of the call is given by T'[ S'[3] ... ( S'[3] + S'[4] - 1 ) ]
(Thus the number of bytes of the transaction is given by S'[4].)
The output data of the call is given by T[ S'[5] ... ( S'[5] + MIN( S'[6], S[0] ) - 1 ) ]
If 0 gas is specified, transfer all gas from the caller to callee gas balance, otherwise transfer only the amount given. If there isn't enough gas in the caller balance, operation fails.
If the value is less than the amount in the caller's balance then nothing is executed and the S'[2] gas gets refunded to the caller.
See Appendix A for execution.
Add any remaining callee gas to the caller's gas.
S[0] = R
where R = 1 when the instrinsic return code of the call is true, R = 0 otherwise.
NOT YET: POST -5 +1
Registers for delayed execution a call where:
The recipient is given by S'[0], when interpreted as an address.
The value is given by S'[1]
The gas to supply the transaction with is given by S'[2] (paid for from the current gas balance)
The input data of the call is given by T'[ S'[3] ... ( S'[3] + S'[4] - 1 ) ]
(Thus the number of bytes of the transaction is given by S'[4].)
Contract pays for itself to run at a defered time from its own GAS supply. The miner has no choice but to execute.
NOT YET: ALARM -6 +1
Registers for delayed execution a call where:
The recipient is given by S'[0], when interpreted as an address.
The value is given by S'[1]
The gas (to convert from ETH at the later time) is given by S'[2]
The number of blocks to wait before executing is S'[3]
The input data of the call is given by T'[ S'[4] ... ( S'[4] + S'[5] - 1 ) ]
(Thus the number of bytes of the transaction is given by S'[5].)
Total gas used now is S'[3] * S'[5] * deferFee.
Contract pays for itself to run at the defered time converting given amount of gas from its ETH balance; if it cannot pay it terminates as a bad transaction. TODO: include baseFee and allow miner freedom to determine whether to execute or not. If not, the next miner will have the chance.
0xf2: RETURN -2 +0
Halts execution.
R := T[ S'[0] ... ( S'[0] + S'[1] - 1 ) ]
Where the output data of the call is specified as R.
Any gas left over gets returned to caller (or in the case of the top-level call, the sender converted back to ETH).
0xff: SUICIDE -1 +0
Halts execution.
FOR ALL i: IF P[i] NOT EQUAL TO 0 THEN B[ S'[0] ] := B[ S'[0] ] + memoryFee
B[ S'[0] ] := B[ S'[0] ] + B[ ADDRESS ]
Removes all contract-related information from the Ethereum system.

>>>>

\documentclass[9pt,oneside]{amsart}
%\usepackage{tweaklist}
\usepackage{url}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[a4paper,width=170mm,top=18mm,bottom=22mm,includeheadfoot]{geometry}
\usepackage{booktabs}
\usepackage{array}
\usepackage{verbatim}
\usepackage{caption}
\usepackage{natbib}
\usepackage{float}
\usepackage{pdflscape}
\usepackage{mathtools}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{afterpage}

\definecolor{lightyellow}{rgb}{1,0.98,0.9}

@@

\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger}
\author{
    Dr. Gavin Wood\\
    CTO, Ethereum Project\\
    gavin@ethereum.org
}
\begin{document}

\pagecolor{lightyellow}

@@

\begin{equation}
H_d \equiv  \begin{cases}
2^{22} & \quad \text{if genesis block}\\
H'_d + \frac{H'_d}{1024} & \quad \text{if $H_t < H'_t + 42$}\\
H'_d - \frac{H'_d}{1024} & \quad \text{otherwise}\\
\end{cases}
\end{equation}

@@

\begin{eqnarray}
a' & \equiv & \mathcal{L}_{160}(\mathcal{S}(\mathcal{R}([s, o])))\\
a & \equiv & \arg \min_x : x \geq a' \wedge \mathbb{S}[x] = \varnothing
\end{eqnarray}

where $\mathcal{S}$ is the SHA3 256-bit hash function, $\mathcal{R}$ is the RLP encoding function, $\mathcal{L}_y(X)$ evaluates to the leftmost $y$ bits of the binary data $X$ and $\mathbb{S}[x]$ is the address state of $x$ or $\varnothing$ if none exists.

@@

\begin{equation}
\mathbb{S}'[r] \equiv ( 0, v, \varnothing, \mathcal{S}(\mathbf{b}) )
\end{equation}

@@

\begin{equation}
\mathbb{S}'[r]_{balance} \equiv \mathbb{S}[r]_{balance} + v
\end{equation}

@@

\begin{itemize}
\item $I_a$, the address of the account which owns the code that is executing.
\item $I_o$, the sender address of the transaction that originated this execution.
\item $I_p$, the price of gas in the transaction that originated this execution.
\item $I_D$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
\item $I_v$, the value passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
\item $I_H$, the block header of the previous block.
\item $I_n$, the block number of the current block.
\item $I_C$, the byte array that is the machine code to be executed.
\end{itemize}

@@

However, instructions do typically alter one or several components of these values. Altered components, on an listed by instruction are noted in Appendix \ref{app:vm}.

@@

Two directions exist for ASIC resistance; firstly make it sequential memory-hard, i.e. engineer the function such that the determination of the nonce requires a lot of memory and that the memory cannot be used in parallel to discover multiple nonces simultaneously. The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.

@@

\begin{equation}
\mathcal{P}(H, n) \equiv \mathcal{B}(\mathcal{S}(\mathcal{S}(\mathcal{R}(H)) \circ n))
\end{equation}

where:
$\mathcal{R}(H)$ is the RLP encoding of the $H$;
$\mathcal{S}$ is the SHA3 hash function accepting an arbitrary length series of bytes and evaluating to a series of 32 bytes (i.e. 256-bit);
$n$ is the nonce, a series of 32 bytes;
$\circ$ is the series concatenation operator;
$\mathcal{B}(X)$ evaluates to the value equal to X when interpreted as a big-endian-encoded integer.

@@

For the release series, we use a more complex proof-of-work. This has yet to be formally defined, but involves two components; firstly that it concerns contract processing, and by extension the evaluation of programs on the EVM. Secondly that it concerns the utilisation of either the block-chain or the full state trie.

As an overview, the output of the function is based upon the system state, defined as the hash of the root node of the state trie. A set of transactions, pseudo-randomly determined from the nonce value and selected from the last $N$ blocks is taken. $N$ is large enough and the selection criteria are such that execution of the transactions requires some non-negligable amount of processing by the EVM. Whenever code is executed on the EVM, it is pseudo-randomly (seeded again by the nonce) corrupted before alteration. Corruption could involve switching addresses with other transactions or rotating them through in the state trie (perhaps to the next address with the same order of magnitude of funds), rotating through instructions that have equivalent stack behaviour (e.g. swapping {\small ADD} for {\small SUB} or {\small GT} for {\small EQ}), or more destructive techniques such as randomly changing opcodes. This results in a problem that both require generalised computation hardware and is sequentially memory (and perhaps even disk) hard.

Any specialised hardware to perform this task could also be leveraged to speed up (and thus drive down costs) of general Ethereum transaction processing.

@@

@@

This paper has introduced, discussed and formally defined the protocol of Ethereum.

@@

\begin{tabular*}{\columnwidth}[h]{rll}
\toprule
Name & Value & Description* \\
\midrule
$G_{step}$ & 1 & Default amount of gas to pay for execution cycle. \\
$G_{stop}$ & 0 & Nothing paid for the {\small STOP} operation. \\
$G_{sha3}$ & 20 & Paid for a {\small SHA3} operation. \\
$G_{sload}$ & 20 & Paid for a {\small SLOAD} operation. \\
$G_{sstore}$ & 100 & Paid for a normal {\small SSTORE} operation (doubled of waived sometimes). \\
$G_{balance}$ & 20 & Paid for a {\small BALANCE} operation. \\
$G_{create}$ & 100 & Paid for a {\small CREATE} operation and for a contract creation transaction. \\
$G_{call}$ & 20 & Paid for a {\small CALL} operation and for a message call transaction. \\
$G_{memory}$ & 1 & Paid for every additional word when expanding memory. \\
$G_{txdata}$ & 5 & Paid for every byte of data or code for a transaction. \\
\bottomrule
\end{tabular*}

@@

As previously specified in section \ref{ch:model}, these definitions take place in the final context there. In particular we assume $O$ is the EVM state-progression function and define the terms pertaining to the next cycle's state $\mathbb{S}'$ $\mathbb{M}'$ such that:

\begin{equation}
O(\mathbb{S}, \mathbb{M}, I) \equiv (\mathbb{S}', \mathbb{M}', I)
\end{equation}

When interpreting 256-bit binary values as integers the representation is big-endian.

When a 32-byte machine datum is converted to and from as a 160-bit address or hash, the leftwards (high-order for BE) 20 bytes are used and the right most 12 are discarded or filled with zeroes.

TODO: Specify minimum memory limits for a general cost definition.

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{0s: Stop and Arithmetic Operations}} \\
\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$.} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
&&&& $\mathbb{M}'_R = []$ \\
\midrule
0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] + \mathbb{M}_T[1]$ \\
\midrule
0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] \times \mathbb{M}_T[1]$ \\
\midrule
0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] - \mathbb{M}_T[1]$ \\
\midrule
0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \lfloor\mathbb{M}_T[0] \div \mathbb{M}_T[1]\rfloor$ \\
\midrule
0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \lfloor\mathbb{M}_T[0] \div \mathbb{M}_T[1]\rfloor$ \\
&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
\midrule
0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] \mod{ \mathbb{M}_T[1] }$ \\
\midrule
0x07 & {\small SMOD} & 2 & 1 & Modulo remainder operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] \mod{ \mathbb{M}_T[1] }$ \\
&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
\midrule
0x08 & {\small EXP} & 2 & 1 & Exponential operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] ^ {\mathbb{M}_T[1] }$ \\
\midrule
0x09 & {\small NEG} & 1 & 1 & Negation operation. \\
&&&& $\mathbb{M}'_T[0] \equiv -\mathbb{M}_T[0]$ \\
&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
\midrule
0x0a & {\small LT} & 2 & 1 & Less-than comparision. \\
&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_T[0] < \mathbb{M}_T[1] \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x0b & {\small GT} & 2 & 1 & Greater-than comparision. \\
&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_T[0] > \mathbb{M}_T[1] \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x0c & {\small EQ} & 2 & 1 & Equality comparision. \\
&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_T[0] = \mathbb{M}_T[1] \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x0d & {\small NOT} & 1 & 1 & Simple not operator. \\
&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_T[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{10s: Bitwise Logic Operations}} \\
\multicolumn{5}{l}{$\mathbb{M}_T[0]_i$ gives the $i$th bit (counting from zero) of $\mathbb{M}_T[0]$} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x10 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_T[0]_i \wedge \mathbb{M}_T[1]_i$ \\
\midrule
0x11 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_T[0]_i \vee \mathbb{M}_T[1]_i$ \\
\midrule
0x12 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_T[0]_i \oplus \mathbb{M}_T[1]_i$ \\
\midrule
0x13 & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
&&&& $\forall i \in [0..7]: \mathbb{M}'_T[0]_i \equiv \begin{cases} \mathbb{M}_T[1]_{(i + 8\mathbb{M}_T[0])} & \text{if} \quad \mathbb{M}_T[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
&&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
\bottomrule
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{20s: SHA3}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x20 & {\small SHA3} & 2 & 1 & Compute SHA3-256 hash. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathcal{S}(\mathbb{M}_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + \mathbb{M}_S[1] - 1) ])$ \\
\bottomrule
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{30s: Environmental Information}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
&&&& $\mathbb{M}'_T[0] \equiv I_a$ \\
\midrule
0x31 & {\small BALANCE} & 0 & 1 & Get balance of currently executing account. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{S}[I_a]_balance$ \\
\midrule
0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
&&&& $\mathbb{M}'_T[0] \equiv I_o$ \\
&&&& This is the sender of original transaction; it's never a contract. \\
\midrule
0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
&&&& $\mathbb{M}'_T[0] \equiv I_c$ \\
&&&& This is the address of the account that is directly responsible for this execution. \\
\midrule
0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
&&&& $\mathbb{M}'_T[0] \equiv I_v$ \\
\midrule
0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathcal{I}_D[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ]$ \\
&&&& This pertains to the input data passed with the message call instruction or transaction. \\
\midrule
0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
&&&& $\mathbb{M}'_T[0] \equiv \lVert \mathcal{I}_D \rVert$ \\
&&&& This pertains to the input data passed with the message call instruction or transaction. \\
\midrule
0x37 & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathcal{I}_p$ \\
&&&& This is gas price specified by the originating transaction.\\
\bottomrule
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{40s: Block Information}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x40 & {\small PREVHASH} & 0 & 1 & Get hash of most recent complete block. \\
&&&& $\mathbb{M}'_T[0] \equiv I_{H_{prevHash}}$ \\
&&&& $I_{H_{prevHash}}$ is the previous block's hash. \\
\midrule
0x41 & {\small COINBASE} & 0 & 1 & Get the block's coinbase address. \\
&&&& $\mathbb{M}'_T[0] \equiv I_{H_{coinbase}}$ \\
\midrule
0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
&&&& $\mathbb{M}'_T[0] \equiv I_{H_{timestamp}}$ \\
\midrule
0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
&&&& $\mathbb{M}'_T[0] \equiv I_{H_{number}}$ \\
&&&& $I_{H_{number}}$ is the total number of ancestor blocks of the current block. \\
\midrule
0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
&&&& $\mathbb{M}'_T[0] \equiv I_{H_{difficulty}}$ \\
\midrule
0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
&&&& $\mathbb{M}'_T[0] \equiv 10^6$ \\
&&&& Presently, the gas limit is hard-coded in the protocol. \\
\bottomrule
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{50s: Stack, Memory, Storage and Flow Operations}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x50 & {\small POP} & 1 & 0 & Remove item from stack. \\
\midrule
0x51 & {\small DUP} & 1 & 2 & Duplicate stack item. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0]$ \\
\midrule
0x52 & {\small SWAP} & 2 & 2 & Exchange stack items. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[1]$ \\
&&&& $\mathbb{M}'_T[1] \equiv \mathbb{M}_T[0]$ \\
\midrule
0x53 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_M[\mathbb{M}_T[0] \dots (\mathbb{M}_T[0] + 31) ]$ \\
\midrule
0x54 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
&&&& $\mathbb{M}'_M[ \mathbb{M}_T[0] \dots (\mathbb{M}_T[0] + 31) ] \equiv \mathbb{M}_T[1]$ \\
\midrule
0x55 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
&&&& $\mathbb{M}'_M[ \mathbb{M}_T[0] ] \equiv (\mathbb{M}_T[1] \mod 256) $ \\
\midrule
0x56 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[\mathbb{M}_T[0]]$ \\
\midrule
0x57 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
&&&& $\mathbb{M}'_S[ \mathbb{M}_T[0] ] \equiv \mathbb{M}_T[1] $ \\
\midrule
0x58 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
&&&& $J(\mathbb{M}) \equiv \mathbb{M}_T[0] $ \\
&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$ \\
\midrule
0x59 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
&&&& $J(\mathbb{M}) \equiv \begin{cases} \mathbb{M}_T[0] & \text{if} \quad \mathbb{M}_T[0] = 0 \\ \mathbb{M}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$ \\
\midrule
0x5a & {\small PC} & 0 & 1 & Get the program counter. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_{pc}$ \\
\midrule
0x5b & {\small MSIZE} & 0 & 1 & Get the size of active memory. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_{i}$ \\
\midrule
0x5c & {\small GAS} & 0 & 1 & Get the amount of available gas. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_{g}$ \\
\bottomrule
\end{tabular*}
